---
output:
  word_document: default
  html_document: default
---

```{r}
cereals_data <- read.csv("C:/Users/rvssu/Downloads/Project2/80-cereals/cereal.csv")
View(cereals_data)
```


```{r}
cereals_data[!complete.cases(cereals_data),]
```


```{r}
head(cereals_data)
```

```{r}
colnames(cereals_data) <- c("Name", "Manufacturer", "Type", "Calories", "Protein", "Fat", "Sodium", "Fibre", "Carbohydrates", "Sugar", "Potassium", "Vitamins", "Shelf", "Weight", "Cups", "Rating")

cereals_data$Manufacturer_Name <- cereals_data$Manufacturer
cereals_data$Type <- gsub("H", "Hot", x = cereals_data$Type)
cereals_data$Type <- gsub("C", "Cold", x = cereals_data$Type)
cereals_data$Manufacturer_Name <- gsub(pattern = "P", replacement = "Post", x = cereals_data$Manufacturer_Name)
cereals_data$Manufacturer_Name <- gsub(pattern = "A", replacement = "American Home Food Products", x = cereals_data$Manufacturer_Name)
cereals_data$Manufacturer_Name <- gsub(pattern = "G", replacement = "General Mills", x = cereals_data$Manufacturer_Name)
cereals_data$Manufacturer_Name <- gsub(pattern = "K", replacement = "Kellogs", x = cereals_data$Manufacturer_Name)
cereals_data$Manufacturer_Name <- gsub(pattern = "N", replacement = "Nabisco", x = cereals_data$Manufacturer_Name)
cereals_data$Manufacturer_Name <- gsub(pattern = "Q", replacement = "Quaker Oats", x = cereals_data$Manufacturer_Name)
cereals_data$Manufacturer_Name <- gsub(pattern = "R", replacement = "Ralston Purina", x = cereals_data$Manufacturer_Name)
```

```{r}
cereal = cereals_data
```

```{r}
View(cereal)
```

```{r}
# Change cereal type and shelf from character to factor
cereal$Type <- as.factor(cereal$Type)
cereal$Shelfs <- as.factor(as.character(cereal$Shelf, levels = c("1","2","3"),labels("1","2","3")))
cereal$Manufacturer <- as.factor(cereal$Manufacturer_Name)
```


```{r}
library(dplyr)

cereal <- cereal %>% mutate(AverageGroup = case_when(Rating <= 35 ~ '3',
                                             Rating > 35  & Rating <= 54 ~ '2',
                                             Rating > 54 ~ '1'))
```
```{r}
cereal$AverageGroup <- as.factor(as.character(cereal$AverageGroup, levels = c("1","2","3"),labels("1","2","3")))
```

# summary

```{r}
conti = subset(cereal, select = -c(Name, Manufacturer_Name,Manufacturer,Shelf, Shelfs,Type,AverageGroup))

summary(conti)
```

```{r}
cate = subset(cereal, select = c(Type,Shelfs,Manufacturer,AverageGroup))
summary(cate)
```


# regression

```{r}
mod1 <- lm(Rating ~ Fibre + Sugar + Calories + Sodium,data = cereal)
#mod1 <- lm(rating ~ mfr + type + sugars + fiber + calories,data = cereals_data)

summary(mod1)
```

Each of the variables is assumed,and there is 0.9 for R-squared,which can be use to predict some scores for other cereals.

let us check distribution for target variable ratings,

# ratings

```{r}
hist(cereal$Rating,col="#3090C7", main = "Distribution of Cereal Ratings", xlab = "Cereal Rating") 
```

# Correlation

```{r}
library(corrplot)
corDat <- subset(cereal, select = -c(Manufacturer_Name,Type,Name,Manufacturer,Cups,Weight,Shelfs,AverageGroup))
corrplot(cor(corDat), method= "circle")
```


```{r}
library(ggcorrplot)
ggcorrplot(cor(corDat), hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="square", 
           colors = c("blue", "white", "green"), 
           title="Correlation", 
           ggtheme=theme_classic)
```

Sugars has been negatively correlated rating.


Let’s see if sugar, negatively correlated with rating, has also an influence on the shelving.

# ecdf

```{r}

```

```{r}
#"Name", "Manufacturer", "Type", "Calories", "Protein", "Fat", "Sodium", "Fibre", "Carbohydrates", "Sugar", "Potassium", "Vitamins", "Shelf", "Weight", "Cups", "Rating"
```

```{r}
bp1 <- subset(cereal, select = Sugar, subset = AverageGroup == "1", drop = T)
bp2 <- subset(cereal, select = Sugar, subset = AverageGroup == "2", drop = T)
bp3 <- subset(cereal, select = Sugar, subset = AverageGroup == "3", drop = T)
```


```{r}
plot.ecdf(bp1, xlab = "Sugar content")
plot.ecdf(bp2, col = "blue",pch = 2, add = TRUE)
plot.ecdf(bp3, col = "red",pch = 5, add = TRUE)
legend(-0.3, 1, legend = c("Rating Group 1", "Rating Group 2", "Rating Group 3"), col = c("black", "blue", "Red"), pch = c(19, 2,5))
#abline(v = 250, lty = 0.6)
```

We can see shelf 2 contains more sugars

Performing boxplot to understand average ratings of shelf

```{r}
# cereal$avggrp = if(cereal$Rating <= 35)   '3'
# cereal$avggrp = (cereal$Rating>35 & cereal$Rating<=54 <- 2
# cereal$avggrp = (cereal$Rating>54 & data$Rating<=29) <- 1
```


# Boxplot

```{r}
boxplot(Sugar~AverageGroup,data = cereal, main="Box plot to understand the Sugar content of Rating group",col = c("green", "red","yellow"))
```

The second row of cereal on the shelf had lower scores showing that customers are not interested in sugars.

manufacturer vs rating

```{r}
library(ggplot2)
ggplot(cereal, aes(Manufacturer,fill = AverageGroup),ggtheme=theme_classic)+
  geom_bar(position = 'dodge')+
  theme_bw()+
  theme(axis.text.y = element_text(angle = 30, hjust = 0.75))+
  coord_flip()+
    scale_y_continuous(breaks=c(2,4,6,8,10,12,14))
#ggplot(data=cereal,aes(Manufacturer_Name,y = ..Count..))
#geom_col(aes(group = Shelfs),stat_identity(position = "dodge"))
```

Let us concentrate on count values.

Most of manufacturers have placed more number of cereals on 3rd shelf. General Mills = 9, Kellogs = 12, Post = 6 and Quaker = 4.

Exception is Nabisco - it has done opposite - More on shelf 1 and least on shelf 3.

The neutral one is Ralston - 4 on shelf 1 and 4 on shelf 3.

# QQ plot Normality

On performing normality on all the variables, Sodium seems to fit the normality graph,

```{r}
a = cereal$Sodium
qqnorm(a)
qqline(a)
```


```{r}
mean(a)
sd(a)
```

```{r}
x <-(a-159.6753)/83.8323 
 plot.ecdf(x)             # adjust x sample mean and standard deviation 
curve(pnorm(x), col = "blue", add = TRUE) # impose normal cdf
```

# K means clustering

```{r}
kmeans_data = corDat
kmeans_data <- kmeans_data[complete.cases(kmeans_data), ]

```
```{r}
wss <- 0
for (i in 1:10) {
  kmeans.output <- kmeans(x = kmeans_data[], centers = i, nstart = 20)
  wss[i] <- kmeans.output$tot.withinss
}
```

```{r}
library(dplyr)
library(tidyr)
data.frame(Clusters = 1:10, WSS = wss) %>% 
  ggplot(aes(x = Clusters, y = WSS)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(name = "Number of Clusters", breaks = 1:10, minor_breaks = NULL) +
  scale_y_continuous(name = "Within Groups Sum of Squares") +
  theme_minimal() #+
  #labs(title = "Scree Plot")
```


```{r}
kmeans.output <- kmeans(x = kmeans_data, centers =3 , nstart = 20)

kmeans.output
```


```{r}
dune_dist <- dist(corDat)
cmd <- cmdscale(dune_dist)
```

```{r}
library(vegan3d)
groups <- levels(factor(kmeans.output$cluster))
ordiplot(cmd, type = "n")
cols <- c("steelblue", "red", "green", "orange")
for(i in seq_along(groups)){
  points(cmd[factor(kmeans.output$cluster) == groups[i], ], col = cols[i], pch = 16)
}
ordispider(cmd, factor(kmeans.output$cluster), label = TRUE)
ordihull(cmd, factor(kmeans.output$cluster), lty = "dotted")

```

```{r}
dd <- cbind(cereal, cluster = kmeans.output$cluster)
```

```{r}
dd$Cluster = as.factor(dd$cluster)
```

# manufacturer vs cluster
```{r}
library(ggplot2)
ggplot(dd, aes(Cluster,fill = Manufacturer))+
  geom_bar(position = 'dodge')+
  theme_minimal()
```
